Overview and Summary of the First Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE1) \note{\scriptsize Authors, please add/check middle initials and affiliations}

### 2014-04-15 WSSSPE1_report/WSSSPE1_summary_report.tex
Add non-breaking spaces and other minor changes.

The First Workshop on Sustainable Software for Science: Practice and
Experiences (WSSSPE1,
\url{http://wssspe.researchcomputing.org.uk/WSSSPE1}) was held on
Sunday, 17 November 2013, in conjunction with the 2013 International
Conference for High Performance Computing, Networking, Storage and
Analysis (SC13, \url{http://sc13.supercomputing.org}).

Because progress in scientific research is dependent on the quality of and
accessibility to software at all levels, it is now critical to address many
challenges related to the development, deployment, and maintenance of reusable
software.
%\note{Ketan: What are the levels in scientific process where software
%is required. It might be interesting to outline these stages and types of
%software required in each stage.}
In addition, it is essential that scientists,
researchers, and students are able to learn and adopt software-related skills
and methodologies. Established researchers are already acquiring some of these
skills, and in particular a specialized class of software developers is
emerging in academic environments as an integral part of
successful research teams. The WSSSPE workshop  provided a forum
for discussion of the challenges around scientific software, including both positions and experiences. The
workshop short papers and debates have been archived to provide a basis for continued
discussion, and have fed into the collaborative writing of this
document. Additional papers based on extended versions of workshop
submissions have been developed as part of this special journal edition. The level of interest in the workshop which involved 
an estimated 90 to 150 participants during the day has led to further activities:
Two additional WSSSPE workshops are planned  (at the 2014 SciPy and SC14 conferences), 
funds to support the workshops have been obtained from 
the National Science Foundation and the Gordon and Betty Moore
Foundation, and the
original workshop website has been turned into a community website for
further discussion and progress. Additionally, a minisymposium  to
further explore some of the key issues raised in WSSSPE is co-organized by a
WSSSPE1 participant at the 2014 Society for Industrial and Applied Mathematics
(SIAM) Annual Meeting on ``Reliable
Computational Science'' (SIAM AN14, \url{http://meetings.siam.org}).

@@

WSSSPE1 was organized in collaboration of the relatively small group
of five organizers and a larger peer-review committee with 36
members. This committee had early influence on the organization, e.g.,
on the specific call for papers.

@@

A peer review process followed the submissions, where
the 58 papers received 181 reviews, an average of 3.12 reviews per
paper. Reviews were completed using a Google form, which allowed
reviewers to choose papers they wanted to review, then to provide
general comments and scores on relevance to the organizers and
to the authors. The review reports enabled the organizers to decide
which papers to associate with the workshop, and allowed the authors
to improve their papers.

@@

Bourne's  presentation was based on his own software experiences.  He emphasized that
sustainability for software ``does not just mean more money from
Government'' (see also Section~\ref{sec:defining-sustainability}).  Other
factors to consider, he mentioned, include costs of production, ease
of maintenance, community involvement, and distribution channels.

>>>>

The First Workshop on Sustainable Software for Science: Practice and
Experiences (WSSSPE1,
\url{http://wssspe.researchcomputing.org.uk/WSSSPE1}) was held on
Sunday, 17 November~2013, in conjunction with the~2013 International
Conference for High Performance Computing, Networking, Storage and
Analysis (SC13, \url{http://sc13.supercomputing.org}).

Because progress in scientific research is dependent on the quality of and
accessibility to software at all levels, it is now critical to address many
challenges related to the development, deployment, and maintenance of reusable
software.
%\note{Ketan: What are the levels in scientific process where software
%is required. It might be interesting to outline these stages and types of
%software required in each stage.}
In addition, it is essential that scientists,
researchers, and students are able to learn and adopt software-related skills
and methodologies. Established researchers are already acquiring some of these
skills, and in particular a specialized class of software developers is
emerging in academic environments as an integral part of
successful research teams. The WSSSPE workshop  provided a forum
for discussion of the challenges around scientific software, including both positions and experiences. The
workshop short papers and debates have been archived to provide a basis for continued
discussion, and have fed into the collaborative writing of this
document. Additional papers based on extended versions of workshop
submissions have been developed as part of this special journal edition. The level of interest in the workshop which involved 
an estimated~90 to~150 participants during the day has led to further activities:
Two additional WSSSPE workshops are planned  (at the~2014 SciPy and SC14 conferences), 
funds to support the workshops have been obtained from 
the National Science Foundation and the Gordon and Betty Moore
Foundation, and the
original workshop website has been turned into a community website for
further discussion and progress. Additionally, a minisymposium  to
further explore some of the key issues raised in WSSSPE is co-organized by a
WSSSPE1 participant at the~2014 Society for Industrial and Applied Mathematics
(SIAM) Annual Meeting on ``Reliable
Computational Science'' (SIAM AN14, \url{http://meetings.siam.org}).

@@

WSSSPE1 was organized in collaboration of the relatively small group
of five organizers and a larger peer-review committee with~36
members. This committee had early influence on the organization, e.g.,
on the specific call for papers.

@@

A peer review process followed the submissions, where
the~58 papers received~181 reviews, an average of~3.12 reviews per
paper. Reviews were completed using a Google form, which allowed
reviewers to choose papers they wanted to review, then to provide
general comments and scores on relevance to the organizers and
to the authors. The review reports enabled the organizers to decide
which papers to associate with the workshop, and allowed the authors
to improve their papers.

@@

Bourne's  presentation was based on his own software experiences.  He emphasized that
sustainability for software ``does not just mean more money from
Government'' (see also Section~\ref{sec:defining-sustainability}).  Other
factors to consider, he mentioned, encompass costs of production, ease
of maintenance, community involvement, and distribution channels.
